The perceptual dimensions of \\ sonority-driven epenthesis

### 2013-09-11 paper.tex
CCC section almost complete except for diagrams

In idiolect \ref{idiolecta}, all coda clusters with a {\sc Sonority Angle} smaller than 1.5 -- the rising and level sonority clusters -- are broken up by epenthesis, while those with a {\sc Sonority Angle} greater than 1.5 -- the falling sonority clusters -- are left intact.

%TODO: diagram

@@

\ex. {\sc Ident(SonAngle)}$<$1.9 $\gg$ {\sc *ComplexCoda}

@@

@@

In the default idiolect, epenthesis occurs between C2 and C3 to form [CC\textipa{1}C].
Only when C1-C2 constitutes a sonority rise or plateau, and C2-C3 constitutes a sonority fall, does epenthesis occur between C1 and C2 to form [C\textipa{1}CC]. Forms and their associated sonority angles are shown in the table below. Underlined forms have alternate forms in other idiolects.

% TODO: table

In this idiolect, all sonority rises and plateaus --- that is, all clusters with a {\sc Sonority Angle} less than 1.5 --- are conflated together. Similarly, all clusters with a {\sc Sonority Angle} greater than 1.5 are conflated together, with the default preference being to epenthesise on the right.

Borrowing a constraint from \citet{rose.2000}, I use {\sc Align-Left} as the markedness constraint that prefers [CC\textipa{1}C] over [C\textipa{1}CC].

@@

@@

\ex. {\sc Ident(SonAngle)}$<$1.5 $\gg$ {\sc Align-Left}
     \a. Fall-\{Rise, Plateau\}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-wzf-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{j@wz1fo} & 1.33 & \cellcolor{lightgray} \\
\hline b. & & \textipa{j@wizfo} & 2.03 $\ast$! & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \b. \{Rise, Plateau\}-\{Rise, Plateau\}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-gdf-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{j@gd1fo} & 1.37 & \\
\hline b. & & \textipa{j@g1dfo} & 0.59 & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \c. Fall-Fall
\begin{center} \renewcommand*\arraystretch{1.2}%h%
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j-a-mst-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{jams1to} & 2.11 $\ast$ & \\
\hline b. & & \textipa{jam1sto} & 2.03 $\ast$ & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \d. \{Rise, Plateau\}-Fall
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-kft-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & & \textipa{j@kfito} & 2.11 $\ast$!  & \cellcolor{lightgray} \\
\hline b. & \ding{43} & \textipa{j@kifto} & 0.59 & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

In a second idiolect discussed in Section 5.2.2 of \citep{rose.2000}, clusters behave much as in the default, except that when C2 is [r] and C3 is sonorant, epenthesis occurs on the right instead.

@@

\ex. \a. Behaviour of clusters with C2=/r/ and C3=sonorant under idiolect 1:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-krm-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & & \textipa{j@krimo} & 1.89 $\ast$! &  \\
\hline b. & \ding{43} & \textipa{j@kirmo} & 0.12 & $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}		 
     \b. Behaviour of clusters with C2=/r/ and C3=sonorant under idiolect 2:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-krm-o}/} & {\sc Ident(SonAngle)}$<$1.9 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{j@krimo} & 1.89 & \\
\hline b. & & \textipa{j@kirmo} & 0.12 & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

In the third idiolect, discussed in Section 5.2 of \citep{rose.2000}, most clusters epenthesise as [C\textipa{1}CC]. This includes clusters

\ex. TODO

In other words, the default preference is to epenthesise as [C\textipa{1}CC], but this is overridden when C2-C3 constitutes a sonority rise or plateau, and C1-C2 constitutes a sonority fall.

@@

\ex. Anchor-L-Root: An element at the left edge of the root corresponds to an element at the left edge of a heavy syllable. \citep[(45)]{rose.2000}

@@

\ex. {\sc Ident(SonAngle)}$<$1.5 $\gg$ {\sc Anchor-L-Root}

% TODO: tableaux

There still remain two clusters whose alternative behaviour remains unexplained according to the rankings we have obtained so far.

\ex. wizf, sigd % TODO

These two datapoints can be explained by retaining the constraint ranking of idiolect 3, but moving the dividing line

@@

\ex. {\sc Ident(SonAngle)}$<$1.0 $\gg$ {\sc Anchor-L-Root}

@@

\ex. \a. Idiolect 1: (sonority rises and plateaus) vs (sonority falls)
     \b. Idiolect 2: (sonority rises and plateaus and RN) vs (other sonority falls)
     \c. Idiolect 3: (sonority rises and plateaus) vs (sonority falls)
     \d. Idiolect 4: (sonority rises) vs (sonority plateaus and falls)

% TODO: diagram showing the various cut-off points.

The {\sc Sonority Rise} metric can achieve the same linear separation for idiolects 1, 3 and 4, but notice that this is not possible for idiolect 2, because % TODO

@@

Variable placcement of these two constraints with respect to {\sc Align-Left} and {\sc Anchor-Left Root}, which themselves can be variably ranked, predict three of the four idiolects.

\ex. Rose's constraint rankings for idiolects 1, 3 and 4.
    \a. Idiolect 1: {\sc SyllCon} $\gg$ {\sc Dep} $\gg$ {\sc Align-L} $\gg$ {\sc Anchor-L Root}
        \begin{itemize}
          \item [CC\textipa{1}C] unless C1-C2 violates {\sc SyllCon}
                (is a sonority rise or plateau) and C2-C3 would not.
        \end{itemize}
     \b. Idiolect 3: {\sc SyllCon} $\gg$ {\sc Dep} $\gg$ {\sc Anchor-L Root} $\gg$ {\sc Align-L} 
        \begin{itemize}
          \item [C\textipa{1}CC] unless C2-C3 violates {\sc SyllCon}
                (is a sonority rise or plateau) and C1-C2 would not.
        \end{itemize}
     \c. Idiolect 4: {\sc SyllCon(Loose)} $\gg$ {\sc Dep} $\gg$ {\sc Anchor-L Root} $\gg$ {\sc Align-L} $\gg$ {\sc SyllCon}
        \begin{itemize}
          \item [C\textipa{1}CC] unless C2-C3 violates {\sc SyllCon(Loose)} 
                (is a sonority rise) and C1-C2 would not.
        \end{itemize}

Thus far, the logic has been very similar to the {\sc Sonority Angle} analysis. Explaining the special behaviour of RN clusters in idiolect 2, requires invoking the {\sc *R-Son} constraint used in the analysis of idiolect ?? of coda clusters.

>>>>

In idiolect \ref{idiolecta}, all coda clusters of rising and level sonority -- that is, those with a {\sc Sonority Angle} smaller than 1.5 -- are broken up by epenthesis, while falling sonority clusters -- those with a {\sc Sonority Angle} greater than 1.5 -- are left intact.

@@

%\vspace{-3em}
%\noindent \resizebox{\linewidth}{!}{\usebox{\sonorityanglehierarchydividingline1.9}}

@@

There are four idiolects to be explained:

\ex. \a. ``Default idiolect'', discussed in Section 4.3 of \citep{rose.2000}. \label{defaultCCCidiolect}
           \a. [C\textipa{1}CC] iff C1-C2 is a fall and C2-C3 is a rise or plateau.
           \b. [CC\textipa{1}C] otherwise.
           \z.
     \b. [CR\textipa{1}N] idiolect, discussed in Section 5.2.2 of \citep{rose.2000}. \label{CRNidiolect}
           \a. [CC\textipa{1}C] if C2 is [r] and C3 is a sonorant.
           \b. Otherwise as default.
           \z.
     \c. [C\textipa{1}CC] idiolect, discussed in Section 5.2.1 of \citep{rose.2000}. \label{C1CCidiolect}
           \a. [CC\textipa{1}C] iff C2-C3 is a fall and C1-C2 is a rise or plateau.
           \b. [C\textipa{1}CC] otherwise.
           \z.
     \d. [\textipa{w1zf}, \textipa{s1gd}] idiolect, discussed in Section 5.2.1 of \citep{rose.2000}. \label{wizfidiolect}
          \a. In this idiolect alone, /wzf/ and /sgd/ epenthesise as [\textipa{w1zf}] and  [\textipa{s1gd}].

@@

\noindent {\bf Default idiolect.} In the default idiolect \ref{defaultCCCidiolect}, epenthesis occurs between C2 and C3 to form [CC\textipa{1}C].
Only when C1-C2 constitutes a sonority rise or plateau, and C2-C3 constitutes a sonority fall, does epenthesis occur between C1 and C2 to form [C\textipa{1}CC]. Forms and their associated sonority angles are shown in the table below. Underlined forms have alternate forms in other idiolects.

\ex. Default idiolect

\begin{tabular}{ccccc}
	& Underlying form      & Surface form & {\sc SonAngle}(C1,C2) & {\sc SonAngle}(C2,C3) \\ \hline
    \multicolumn{5}{c}{Fall-\{Rise, Plateau\}: [CC\textipa{1}C]} \\ \hline
	a. & \textipa{j@-rk't'-o} & \textipa{j@nk'1t'o} & 2.36 & 1.37 \\
	b. & \textipa{j@-rks-o}   & \textipa{j@nk1so}   & 2.36 & 0.59 \\
    c. & \textipa{j@-rk'm-o}  & \textipa{j@nk'1mo}  & 2.36 & 0.27 \\
    d. & \textipa{j@-\underline{wzf}-o} & \textipa{j@wz1fo} & 2.03 & 1.33 \\
	e. & \textipa{j@-mxr-o}   & \textipa{j@mx1ro}   & 2.03 & 0.22 \\
    f. & \textipa{j@-\underline{sgd}-o} & \textipa{j@sg1do}   & 2.11 & 1.37 \\  
    g. & \textipa{j@-sdB-o}   & \textipa{j@sd1Bo}   & 2.11 & 0.18 \\ \hline
    \multicolumn{5}{c}{\{Rise, Plateau\}-\{Rise, Plateau\}: [CC\textipa{1}C]} \\ \hline
    h. & \textipa{j@-\underline{gdf}-o} & \textipa{j@gd1fo} & 1.37 & 0.59 \\
    i. & \textipa{j@-\underline{gdr}-o} & \textipa{j@gd1ro} & 1.37 & 0.12 \\
    j. & \textipa{j@-\underline{kmr}-o} & \textipa{j@km1ro} & 0.27 & 0.46 \\
    k. & \textipa{j@-\underline{sfr}-o} & \textipa{j@sf1ro} & 1.33 & 0.22 \\
    l. & \textipa{j@-\underline{sBr}-o} & \textipa{j@sB1ro} & 0.34 & 0.73 \\ \hline
    \multicolumn{5}{c}{Fall-Fall: [CC\textipa{1}C]} \\ \hline
    m. & \textipa{j@-rfk-o} & \textipa{j@nf1ko} & 2.03 & 2.11 \\ 
    n. & \textipa{j@-rmd-o} & \textipa{j@rm1do} & 1.89 & 2.36 \\ 
    o. & \textipa{j-a-mst-o}& \textipa{jams1to} & 2.03 & 2.11 \\ \hline
    \multicolumn{5}{c}{\{Rise,Plateau\}-Fall: [C\textipa{1}CC]} \\ \hline
    p. & \textipa{j@-kft-o} & \textipa{j@k1fto} & 0.59 & 2.11 \\
    q. & \textipa{j@-dmd-o} & \textipa{j@d1mdo} & 0.27 & 2.36 \\
    r. & \textipa{j@-drs-o} & \textipa{j@d1rso} & 0.12 & 2.21 \\ 
    s. & \textipa{j@-\underline{k'rm}-o} & \textipa{j@k'1rmo} & 0.12 & 1.89 \\ 
    t. & \textipa{j@-\underline{srB}-o}  & \textipa{j@s1rBo}  & 0.22 & 1.57 \\
    u. & \textipa{j@-frt-o} & \textipa{j@f1rto} & 0.22 & 2.36 \\
    v. & \textipa{j@-sBx-o} & \textipa{j@s1Bxo} & 0.34 & 2.17 \\
    w. & \textipa{j-a-mrg-o} & \textipa{jam1rgo} & 0.46 & 2.36 \\ \hline
    \multicolumn{5}{c}{Non-jussive forms} \\ \hline
    x. & \textipa{j-t-Bt'@b@t'} & \textipa{j1t1Bt'@b@t'} & 0.18 & 2.38 \\
    y. & \textipa{j-t-rs@n@s} & \textipa{j1t1rs@n@s} & 0.12 & 2.21 \\
    z. & \textipa{j-t-kB@s@s} & \textipa{j1tk1B@s@s} & 1.37 & 0.12 \\
    A. & \textipa{j-t-k't'@k'@t'} & \textipa{j1tk'1t'@k'@t'} & 1.37 & 1.37 \\
\end{tabular}

The default preference in this idiolect is to epenthesise on the right, between C2 and C3.
I encode this preference with a markedness constraint borrowed from \citep{rose.2000}'s analysis:

@@

The only time that {\sc Align-Left} is overruled is when breaking up the C2-C3 cluster would be too ``expensive'' in faithfulness terms --- when the cluster consists of a sonority fall --- and breaking up the C1-C2 cluster is comparatively cheap, that is, when C1-C2 is a sonority rise or plateau. In {\sc Sonority Angle} terms, {\sc Align-Left} is overruled only when breaking the C2-C3 cluster would entail creating a {\sc Sonority Angle} greater than 1.5, when the alternative of breaking the C1-C2 cluster would create a {\sc Sonority Angle} smaller than 1.5.

@@

\ex. {\sc Ident(SonAngle)}$<$1.5 $\gg$ {\sc Align-Left}
     \a. Fall-\{Rise, Plateau\}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-wzf-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{j@wz1fo} & 1.33 & \cellcolor{lightgray} \\
\hline b. & & \textipa{j@wizfo} & 2.03 $\ast$! & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \b. \{Rise, Plateau\}-\{Rise, Plateau\}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-gdf-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline c. & \ding{43} & \textipa{j@gd1fo} & 1.37 & \\
\hline d. & & \textipa{j@g1dfo} & 0.59 & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \c. Fall-Fall
\begin{center} \renewcommand*\arraystretch{1.2}%h%
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j-a-mst-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline e. & \ding{43} & \textipa{jams1to} & 2.11 $\ast$ & \\
\hline f. & & \textipa{jam1sto} & 2.03 $\ast$ & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \d. \{Rise, Plateau\}-Fall
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-kft-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline g. & & \textipa{j@kfito} & 2.11 $\ast$!  & \cellcolor{lightgray} \\
\hline h. & \ding{43} & \textipa{j@kifto} & 0.59 & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\noindent {\bf Remark on conflation.} Notice that all distinctions among the set of rising and level sonority clusters on the one hand, and the distinctions among the set of falling sonority clusters, are conflated. For instance, consider the cluster /mst/. {\sc SonAngle}(m,s), at 2.03, is smaller than {\sc SonAngle}(s,t), at 2.11, yet epenthesis occurs on the right, creating the larger {\sc Sonority Angle}. This is because the distinction between {\sc Sonority Angles} of 2.03 and 2.11 is irrelevant under the given  constraint ranking, since both 2.03 and 2.11 violate the relevant faithfulness constraint {\sc Ident(SonAngle)}$<$1.5. This is possible because I have adopted the position that the {\sc Ident(SonAngle)}$<n$ constraints can be freely ranked.

Adopting the converse strategy --- that {\sc Ident(SonAngle)}$<n$ constraints have a universal ranking --- would not predict [\textipa{ms1t}]. Instead we would have the following situation:

\ex. Universal ranking cannot derive [\textipa{j@ms1to}]

\vspace{-3em}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-mst-o}/} & {\sc Ident(SonAngle)}$<$2.1 & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a.  &\frownie & \textipa{j@ms1to} & 2.11 $\ast$! & \cellcolor{lightgray}2.11 $\ast$ & \cellcolor{lightgray} \\
\hline b. & \ding{43} & \textipa{j@m1sto} & 2.03 & \cellcolor{lightgray}2.03 $\ast$ & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

\noindent {\bf [CR\textipa{1}N] idiolect.} In a second idiolect \ref{CRNidiolect} (c.f. Section 5.2.2 of \citep{rose.2000}), clusters behave much as in the default, except that when C2 is [r] and C3 is sonorant, epenthesis occurs on the right instead.

@@

\ex. \a. Behaviour of clusters with C2=/r/ and C3=sonorant under idiolect \ref{defaultCCCidiolect}:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-krm-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & & \textipa{j@krimo} & 1.89 $\ast$! &  \\
\hline b. & \ding{43} & \textipa{j@kirmo} & 0.12 & $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}		 
     \b. Behaviour of clusters with C2=/r/ and C3=sonorant under idiolect \ref{CRNidiolect}:
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-krm-o}/} & {\sc Ident(SonAngle)}$<$1.9 & {\sc Align-Left} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{j@krimo} & 1.89 & \\
\hline b. & & \textipa{j@kirmo} & 0.12 & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\noindent {\bf [C\textipa{1}CC] idiolect.} In the third idiolect \ref{C1CCidiolect} (c.f. Section 5.2 of of \citep{rose.2000}), most clusters epenthesise as [C\textipa{1}CC]. This includes clusters

\ex. Default idiolect

\begin{tabular}{ccccc}
	& Underlying form      & Surface form & {\sc SonAngle}(C1,C2) & {\sc SonAngle}(C2,C3) \\ \hline
    \multicolumn{5}{c}{Fall-\{Rise, Plateau\}: [CC\textipa{1}C]} \\ \hline
	a. & \textipa{j@-rk't'-o} & \textipa{j@nk'1t'o} & 2.36 & 1.37 \\
	b. & \textipa{j@-rks-o}   & \textipa{j@nk1so}   & 2.36 & 0.59 \\
    c. & \textipa{j@-rk'm-o}  & \textipa{j@nk'1mo}  & 2.36 & 0.27 \\
    d. & \textipa{j@-\underline{wzf}-o} & \textipa{j@wz1fo} & 2.03 & 1.33 \\
	e. & \textipa{j@-mxr-o}   & \textipa{j@mx1ro}   & 2.03 & 0.22 \\
    f. & \textipa{j@-\underline{sgd}-o} & \textipa{j@sg1do}   & 2.11 & 1.37 \\  
    g. & \textipa{j@-sdB-o}   & \textipa{j@sd1Bo}   & 2.11 & 0.18 \\ \hline
    \multicolumn{5}{c}{\{Rise, Plateau\}-\{Rise, Plateau\}: [C\textipa{1}CC]} \\ \hline
    h. & \textipa{j@-\underline{gdf}-o} & \textipa{j@g1dfo} & 1.37 & 0.59 \\
    i. & \textipa{j@-\underline{gdr}-o} & \textipa{j@g1dro} & 1.37 & 0.12 \\
    j. & \textipa{j@-\underline{kmr}-o} & \textipa{j@k1mro} & 0.27 & 0.46 \\
    k. & \textipa{j@-\underline{sfr}-o} & \textipa{j@s1fro} & 1.33 & 0.22 \\
    l. & \textipa{j@-\underline{sBr}-o} & \textipa{j@s1Bro} & 0.34 & 0.73 \\ \hline
    \multicolumn{5}{c}{Fall-Fall: [CC\textipa{1}C]} \\ \hline
    m. & \textipa{j@-rfk-o} & \textipa{j@nf1ko} & 2.03 & 2.11 \\ 
    n. & \textipa{j@-rmd-o} & \textipa{j@rm1do} & 1.89 & 2.36 \\ 
    o. & \textipa{j-a-mst-o}& \textipa{jams1to} & 2.03 & 2.11 \\ \hline
    \multicolumn{5}{c}{\{Rise,Plateau\}-Fall: [C\textipa{1}CC]} \\ \hline
    p. & \textipa{j@-kft-o} & \textipa{j@k1fto} & 0.59 & 2.11 \\
    q. & \textipa{j@-dmd-o} & \textipa{j@d1mdo} & 0.27 & 2.36 \\
    r. & \textipa{j@-drs-o} & \textipa{j@d1rso} & 0.12 & 2.21 \\ 
    s. & \textipa{j@-\underline{k'rm}-o} & \textipa{j@k'1rmo} & 0.12 & 1.89 \\ 
    t. & \textipa{j@-\underline{srB}-o}  & \textipa{j@s1rBo}  & 0.22 & 1.57 \\
    u. & \textipa{j@-frt-o} & \textipa{j@f1rto} & 0.22 & 2.36 \\
    v. & \textipa{j@-sBx-o} & \textipa{j@s1Bxo} & 0.34 & 2.17 \\
    w. & \textipa{j-a-mrg-o} & \textipa{jam1rgo} & 0.46 & 2.36 \\ \hline
\end{tabular}

In this idiolect, the default preference is to epenthesise as [C\textipa{1}CC], but this is overridden when C2-C3 constitutes a sonority rise or plateau, and C1-C2 constitutes a sonority fall.

@@

\ex. Anchor-Left Root: An element at the left edge of the root corresponds to an element at the left edge of a heavy syllable. \citep[(45)]{rose.2000}

@@

\ex. {\sc Ident(SonAngle)}$<$1.5 $\gg$ {\sc Anchor-Left Root} \label{CCiCidiolectranking}
     \a. Fall-\{Rise, Plateau\}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-wzf-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Anchor-L Rt} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{j@wz1fo} & 1.33 & \cellcolor{lightgray}$\ast$ \\
\hline b. & & \textipa{j@wizfo} & 2.03 $\ast$! & \cellcolor{lightgray} \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \b. \{Rise, Plateau\}-\{Rise, Plateau\}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-gdf-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Anchor-L Rt} \\[0.5ex]
\hline \hline c. & & \textipa{j@gd1fo} & 1.37 & $\ast$! \\
\hline d. & \ding{43}  & \textipa{j@g1dfo} & 0.59 &  \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \c. Fall-Fall
\begin{center} \renewcommand*\arraystretch{1.2}%h%
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j-a-mst-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Anchor-L Rt} \\[0.5ex]
\hline \hline e. & & \textipa{jams1to} & 2.11 $\ast$! & \\
\hline f. & \ding{43} & \textipa{jam1sto} & 2.03 $\ast$ & $\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \d. \{Rise, Plateau\}-Fall
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-kft-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Anchor-L Rt} \\[0.5ex]
\hline \hline g. & & \textipa{j@kf1to} & 2.11 $\ast$!  & \cellcolor{lightgray}$\ast$ \\
\hline h. & \ding{43} & \textipa{j@k1fto} & 0.59 & \cellcolor{lightgray} \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

This ranking predicts that fall-fall clusters should epenthesise on the left. Items (m) and (n) are accounted for as they are subject to the {\sc Jussive-R} constraint, but the /mst/ cluster in (o) is predicted to surface as [m\textipa{1}st], whereas Rose marks it as being epenthesised as [ms\textipa{1}t] without variation. On the other hand, one of Rose's constraint rankings produces this form. Moreover, Degif Petros Banksira (p.c.), a native speaker of Chaha, confirms that he produces the form [jam\textipa{1}sto]. Therefore this prediction actually proves correct.

\noindent {\bf [\textipa{w1zf}, \textipa{s1gd} idiolect.]} There still remain two clusters whose alternative behaviour remains unexplained according to the rankings we have obtained so far.

\ex. \a. /\textipa{j@-wzf-o}/ can surface as [\textipa{j@-w1zf-o}], but this is not predicted by any ranking so far, including \ref{CCiCidiolectranking}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-wzf-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Anchor-L Root} \\[0.5ex]
\hline \hline a. & \frownie & \textipa{j@w1zfo} & 2.03 $\ast$! & \cellcolor{lightgray} \\
\hline b. & \ding{43} & \textipa{j@wz1fo} & 1.32 & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \b. /\textipa{j@-sgd-o}/ can surface as [\textipa{j@-s1gd-o}], but this is not predicted by any ranking so far, including \ref{CCiCidiolectranking}
\begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-sgd-o}/} & {\sc Ident(SonAngle)}$<$1.5 & {\sc Anchor-L Root} \\[0.5ex]
\hline \hline c. & \frownie & \textipa{j@s1gdo} & 2.11 $\ast$! & \cellcolor{lightgray} \\
\hline d. & \ding{43} & \textipa{j@sg1do} & 1.37 & \cellcolor{lightgray}$\ast$ \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

These two datapoints can be explained by retaining the constraint ranking of idiolect 3, but moving the dividing line to 1.0, between sonority rises on the one hand, and sonority plateaus and falls on the other.

@@

\ex. {\sc Ident(SonAngle)}$<$1.0 $\gg$ {\sc Anchor-L-Root}
     \a. \begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-wzf-o}/} & {\sc Ident(SonAngle)}$<$1.0 & {\sc Anchor-L Root} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{j@w1zfo} & 2.03 $\ast$ & \\
\hline b. & & \textipa{j@wz1fo} & 1.32 $\ast$ & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}
     \b. \begin{center} \renewcommand*\arraystretch{1.2}
\scalebox{1}[1]{\begin{tabular}[t]{|rrl||c|c|} \hline 
\multicolumn{3}{|c||}{Input:~/\textipa{j@-sgd-o}/} & {\sc Ident(SonAngle)}$<$1.0 & {\sc Anchor-L Root} \\[0.5ex]
\hline \hline a. & \ding{43} & \textipa{j@s1gdo} & 2.11 $\ast$ & \\
\hline b. & & \textipa{j@sg1do} & 1.37 $\ast$ & $\ast$! \\
\hline \end{tabular}} \renewcommand*\arraystretch{1} \end{center}

@@

\ex. \a. Idiolect \ref{defaultCCCidiolect} (sonority rises and plateaus) vs (sonority falls)
     \b. Idiolect \ref{CRNidiolect}: (sonority rises and plateaus and RN) vs (other sonority falls)
     \c. Idiolect \ref{C1CCidiolect}: (sonority rises and plateaus) vs (sonority falls)
     \d. Idiolect \ref{wizfidiolect}: (sonority rises) vs (sonority plateaus and falls)

% TODO: SonAngle diagram showing the various cut-off points.

The {\sc Sonority Rise} metric can achieve the same linear separation for idiolects 1, 3 and 4, but notice that this is not possible for idiolect \ref{CRNidiolect}, because {\sc Sonority Rise} cannot achieve a linear separation of the sonority rises, plateaus and RN from the remainder of the falling sonority clusters.

% TODO: SonRise diagram

@@

Variable placement of these two constraints with respect to {\sc Align-Left} and {\sc Anchor-Left Root}, which themselves can be variably ranked, predict three of the four idiolects.

\ex. Rose's constraint rankings for idiolects \ref{defaultCCCidiolect}, \ref{C1CCidiolect} and \ref{wizfidiolect}.
    \a. Idiolect \ref{defaultCCCidiolect}: {\sc SyllCon} $\gg$ {\sc Dep} $\gg$ {\sc Align-L} $\gg$ {\sc Anchor-L Root}
          \a. [CC\textipa{1}C] unless C1-C2 violates {\sc SyllCon}
               (is a sonority rise or plateau) and C2-C3 would not.
          \z.
     \b. Idiolect \ref{C1CCidiolect}: {\sc SyllCon} $\gg$ {\sc Dep} $\gg$ {\sc Anchor-L Root} $\gg$ {\sc Align-L} 
          \a. \item [C\textipa{1}CC] unless C2-C3 violates {\sc SyllCon}
                (is a sonority rise or plateau) and C1-C2 would not.
          \z.
     \c. Idiolect \ref{wizfidiolect}: {\sc SyllCon(Loose)} $\gg$ {\sc Dep} $\gg$ {\sc Anchor-L Root} $\gg$ {\sc Align-L} $\gg$ {\sc SyllCon}
          \a. [C\textipa{1}CC] unless C2-C3 violates {\sc SyllCon(Loose)} 
                (is a sonority rise) and C1-C2 would not.
          \z.

Thus far, the logic has been very similar to the {\sc Sonority Angle} analysis. Explaining the special behaviour of RN clusters in the [CR\textipa{1}N] idiolect \ref{CRNidiolect}, requires invoking the {\sc *R-Son} constraint used in the analysis of idiolects \ref{idiolectb} and \ref{idiolectd} of the coda clusters.
